Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-26 NEUROC/sec4-experiments.tex
grammer和检查了sec4.2

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gaan.pdf}}
    \caption{Training time breakdown on the layer level. The training time of each layer included the time spent on the forward, backward and evaluation phases. Each layer was further decomposed into the vertex and the edge calculation stages.}
    \label{fig:exp_vertex_edge_cal_proportion}
\end{figure}

@@

We further investigated the most time-consuming step of the edge calculation stage.
%
In the implementation of PyG, the edge calculation stage consisted of four steps: collection, messaging, aggregation, and vector updating, as shown in \figurename~\ref{fig:steps_in_edge_calculation}.
%
The edge index was a matrix with $|\mathcal{E}|$ rows and two columns.
%
It holded the edge set of the graph.
%
The two columns of the edge index stored the source and the target vertex IDs of each edge.
%
The collection step copied the hidden vectors from the previous GNN layer $\MyVec{h}^l_y$ and $\MyVec{h}^l_x$ to the both endpoints of each edge $e_{y,x}$ in the edge index, forming the parameter tensor $[\MyVec{h}^l_{y}, \MyVec{h}^l_{x}, \MyVec{e}_{y,x}]$ of the messaging function $\phi^l$.
%
This step only involved data movement.
%
The messaging step called the messaging function $\phi^l$ on all edges to get message vectors $\MyVec{m}^l_{y,x}$.
%
The aggregation step aggregated the message vectors with the same target vertex into an aggregated vector $\MyVec{s}^l_x$.
%
The vector updating step was optional.
%
It performed an additional transformation on the aggregated vectors (for example, adding the bias in GCN).

@@

The most time-consuming operators in the four GNNs were the matrix multiplication \texttt{mm} and the elementwise multiplication \texttt{mul}, \emph{making GNN training suitable for GPUs}.
%
Although the aggregation step in the edge calculation stage was relatively simple (like sum and mean), the related operators--\texttt{scatter\_add} and \texttt{gather}--still consumed a certain amount of the time.
%
The two operators had to synchronize between hardware threads to avoid updating the same aggregated vector at the same time.
%
They also conducted non-regular memory access with the access pattern determined by the edge set dynamically.
%
For GPUs, they were less efficient than \texttt{mm}.
%
The index-based selection operator \texttt{index\_select} used in the collection step consumed about 10\% of the training time in all GNNs.
%
Improving the efficiency of \texttt{scatter\_add}/\texttt{gather}/\texttt{index\_select} can benefit all kinds of GNNs.

@@

%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Here, we used GCN as an example. 
%
\figurename~\ref{fig:inference_analysis_on_layer_edge_level_of_gcn} shows the results of GCN inference on the \emph{layer} level and the \emph{step} level.
%
By cross-comparing \figurename~\ref{fig:exp_vertex_edge_cal_proportion}(a), 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}(a), and \figurename~\ref{fig:exp_edge_calc_decomposition}(a),
we could conclude that the time breakdowns were very similar.
%
For the other GNNs, we observed similar phenomena.
%

@@

% 绝对时间对比图
\begin{figure}[H]
    \centering
    \subfloat[\texttt{amc}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-computers.pdf}}
    \subfloat[\texttt{amp}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-photo.pdf}}
    \subfloat[\texttt{fli}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_flickr.pdf}}
    \caption{Wall-clock training/inference time on different datasets.}
    \label{fig:compare_wall_clock_time_of_training_and_inference}
\end{figure}

@@

The top time-consuming basic operators of training and inference also showed a certain degree of difference.
%
Some of the top time-consuming operators during training were replaced by new operators in inference.
%
Figure~\ref{fig:exp_inference_top_basic_ops} in the response compares the top 5 time-consuming basic operators in inference.
%
For GCN, the \texttt{index} operator used in the prediction layer became the new top 5 time-consuming operator, replacing the \texttt{gather} operator used in the backward phase in training.
%
%    The basic operators related to the edge calculation (\texttt{scatter\_add}, \texttt{index\_select}, and \texttt{mul}) still consumed the majority of the inference time.
%
For GGNN, \texttt{index} operator in the prediction layer also became one of the top 5 time-consuming basic operators.
%
%    Due to the high time complexity of the vertex updating function $\gamma$ in GGNN, the basic operators related to the vertex calculation stage (\texttt{mm} and \texttt{thnn\_fused\_gru\_cell}) still consumed near half of the inference time.
%
For GAT, the \texttt{input\_put\_impl} operator (used in the backward phase) in Figure~\ref{fig:exp_top_basic_ops}(c) was replaced by the \texttt{scatter\_add} operator used in the forward edge calculation in Figure~\ref{fig:exp_inference_top_basic_ops}(c).
%
%    The \texttt{scatter\_add} operator was related to the edge calculation step.
%
For GaAN, the \texttt{mm} operator was replaced by the \texttt{cat} operator used in the vertex updating function.

Although some specific time-consuming operators were different, the performance bottlenecks on the operator level were same for training and inference.
%
The matrix multiplication \texttt{mm} and the element-wise multiplication \texttt{mul} operators were still time-consuming for inference, making GNN inference also {suitable for GPUs}.
%
The basic operators related to the collection and aggregation steps in the edge calculation still consumed non-trivial time in inference.

>>>>

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gaan.pdf}}
    \caption{Training time breakdown on the layer level. The training time of each layer included the time spent on the forward, backward, and evaluation phases. Each layer was further decomposed into the vertex and the edge calculation stages.}
    \label{fig:exp_vertex_edge_cal_proportion}
\end{figure}

@@

We further investigated the most time-consuming step of the edge calculation stage.
%
In the implementation of PyG, the edge calculation stage consisted of four steps: collection, messaging, aggregation, and vector updating, as shown in \figurename~\ref{fig:steps_in_edge_calculation}.
%
The edge index was a matrix with $|\mathcal{E}|$ rows and two columns.
%
It held the edge set of the graph.
%
The two columns of the edge index stored the source and the target vertex IDs of each edge.
%
The collection step copied the hidden vectors from the previous GNN layer $\MyVec{h}^l_y$ and $\MyVec{h}^l_x$ to the both endpoints of each edge $e_{y,x}$ in the edge index, forming the parameter tensor $[\MyVec{h}^l_{y}, \MyVec{h}^l_{x}, \MyVec{e}_{y,x}]$ of the messaging function $\phi^l$.
%
This step only involved data movement.
%
The messaging step called the messaging function $\phi^l$ on all edges to get message vectors $\MyVec{m}^l_{y,x}$.
%
The aggregation step aggregated the message vectors with the same target vertex into an aggregated vector $\MyVec{s}^l_x$.
%
The vector updating step was optional.
%
It performed an additional transformation on the aggregated vectors (for example, adding the bias in GCN).

@@

The most time-consuming operators in the four GNNs were the matrix multiplication \texttt{mm} and the elementwise multiplication \texttt{mul}, \emph{making GNN training suitable for GPUs}.
%
Although the aggregation step in the edge calculation stage was relatively simple (like sum and mean), the related operators--\texttt{scatter\_add} and \texttt{gather}--still consumed a certain amount of the time.
%
The two operators had to synchronize between hardware threads to avoid updating the same aggregated vector at the same time.
%
They also conducted non-regular memory access with the access pattern determined by the edge set dynamically.
%
For GPUs, they were less efficient than \texttt{mm}.
%
The index-based selection operator \texttt{index\_select} used in the collection step consumed about 10\% of the training time in all GNNs.
%
Improving the efficiency of \texttt{scatter\_add}/\texttt{gather}/\texttt{index\_select} could benefit all kinds of GNNs.

@@

%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Taking GCN as an example, \figurename~\ref{fig:inference_analysis_on_layer_edge_level_of_gcn} shows the results of GCN inference on the \emph{layer} level and the \emph{step} level.
%
By cross-comparing \figurename~\ref{fig:exp_vertex_edge_cal_proportion}(a), 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}(a), and \figurename~\ref{fig:exp_edge_calc_decomposition}(a),
we could conclude that the time breakdowns were very similar.
%
For the other GNNs, we observed similar phenomena.
%

@@

% 绝对时间对比图
\begin{figure}[H]
    \centering
    \subfloat[\texttt{amc}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-computers.pdf}}
    \subfloat[\texttt{amp}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-photo.pdf}}
    \subfloat[\texttt{fli}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_flickr.pdf}}
    \caption{Wall-clock inference time on different datasets.}
    \label{fig:compare_wall_clock_time_of_training_and_inference}
\end{figure}

@@

The top time-consuming basic operators of training and inference also showed a certain degree of difference.
%
Some of the top time-consuming operators during training were replaced by new operators in inference.
%
Figure~\ref{fig:exp_inference_top_basic_ops} in the response compares the top 5 time-consuming basic operators in inference.
%
For GCN, the \texttt{index} operator used in the prediction layer became the new top 5 time-consuming operators, replacing the \texttt{gather} operator used in the backward phase in training.
%
%    The basic operators related to the edge calculation (\texttt{scatter\_add}, \texttt{index\_select}, and \texttt{mul}) still consumed the majority of the inference time.
%
For GGNN, \texttt{index} operator in the prediction layer also became one of the top 5 time-consuming basic operators.
%
%    Due to the high time complexity of the vertex updating function $\gamma$ in GGNN, the basic operators related to the vertex calculation stage (\texttt{mm} and \texttt{thnn\_fused\_gru\_cell}) still consumed near half of the inference time.
%
For GAT, the \texttt{input\_put\_impl} operator (used in the backward phase) in Figure~\ref{fig:exp_top_basic_ops}(c) was replaced by the \texttt{scatter\_add} operator used in the forward edge calculation in Figure~\ref{fig:exp_inference_top_basic_ops}(c).
%
%    The \texttt{scatter\_add} operator was related to the edge calculation step.
%
For GaAN, the \texttt{mm} operator was replaced by the \texttt{cat} operator used in the vertex updating function.

Although some specific time-consuming operators were different, the performance bottlenecks on the operator level were the same for training and inference.
%
The matrix multiplication \texttt{mm} and the element-wise multiplication \texttt{mul} operators were still time-consuming for inference, making GNN inference also {suitable for GPUs}.
%
The basic operators related to the collection and aggregation steps in the edge calculation still consumed non-trivial time in inference.
